Add Plant Functional Type (PFT) labels

### 2021-04-27 R/add_plant_functional_type.Rmd
updated descriptions. removed automation code attempts that didnt work. need to apply the LUT to add a label to the cleaned spectra

Plant functional types (PFTs) originally proposed: coniferous forest, deciduous forest, woody shrub, grass/herb, and bare ground.

@@

## Use NEON growthForm observations

@@

# summarize the merged growthform data with "tree" and "shrub" growth forms
cleaned_spectra_merged %>% 
  distinct(spectraID, .keep_all = TRUE) %>% 
  # keep just the tree and shrub growth forms, remove rows with "sampling" and "liana"
  filter(growthForm_simple %in% c("tree","shrub")) %>% 
  group_by(genusSpecies, growthForm_simple) %>% 
    tally() %>% 
  # display as a nicely formatted table
  kableExtra::kable() %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", 
                            full_width = F, 
                            position = "left")

@@

- common name
- pft_MRA (based on Macrosystems proposed labels)
    - coniferous_forest
    - deciduous_forest
    - woody_shrub
- pft_MODIS (based on MODIS land cover classes)
    - evergreen-needleleaf_forest
    - evergreen-broadleaf_forest
    - deciduous-needleleaf_forest
    - deciduous-broadleaf_forest
    - shrub
- phenology (evergreen or deciduous)
- leaf type (broadleaf or needleleaf)
- growth form (shrub or tree, based on NEON field based observations)

I searched for each taxon on USDA plants and determined the phenology and leaf type. 
- shrubs that are angiosperms are broadleaf
- group = dicot or monocot, you can say broadleaf
- group = gymnosperm will be needleleaf

Utilized the NEON vst_apparentindividual data for growthform.

@@

@@

```{r install_load_packages, message=FALSE}
#install.packages("request")
#library(devtools)
#devtools::install_github("sckott/request")
library(request)
library(tibble)
library(tidyverse)
library(here)
```

```{r usda_plants}
# Set the base url
root <- api("https://plantsdb.xyz")

root %>% api_path(heartbeat)
```

^^ My results here do not match those that show up on the tutorial page above. Maybe the API is down or the URL has changed?

Based on the GitHub issue discussion (https://github.com/USDA/USDA-APIs/issues/7), someone uploaded USDA plants data here from 2016: https://uofi.app.box.com/v/usdaplants

```{r try_download, eval=FALSE, include=FALSE}
# Specify URL where file is stored 
url <- "https://uofi.app.box.com/v/usdaplants/folder/6688524249/usda_plants_20160713.csv"
destFile <- file.path(here::here(), "analysis","data", "usda_plants_20160713.csv")
# Apply download.file function in R
download.file(url, destFile)
```

```{r, include = FALSE, eval=FALSE}
#usda_plants <- read_csv(file.path(here::here(), "analysis", "data", "usda_plants_20160713.csv"))

# Error in nchar(x) : invalid multibyte string, element 2
# ^ Issue with some of the column names in the .csv? 
```

@@

```{r load_packages_bien, message=FALSE, warning=FALSE}
library(BIEN)
library(ape) # Package for working with phylogenies in R
library(maps) # Useful for making quick maps of occurrences
library(sp) # A package for spatial data
```

BIEN traits and taxonomic data

"Let’s say we’re interested in the genus [Abies], and we’d like to get an idea of how many species there are in this genus and what higher taxa it falls within."

*Abies* genus is fir trees (coniferous trees). The "higher_plant_group" returned by the BIEN taxonomy function is "gymnosperms (conifers)"

```{r bien_taxonomy_1}
Abies_taxonomy<-BIEN_taxonomy_genus(genus = "Abies")

#View(Abies_taxonomy)
Abies_taxonomy[1:3,]
```

*Acer rubrum* species is the red maple (deciduous trees). The "higher_plant_group" returned by the BIEN taxonomy is "flowering plants".

```{r bien_taxonomy_2}
Acer_rubrum_taxonomy <- BIEN_taxonomy_species(species = "Acer rubrum")
Acer_rubrum_taxonomy[1:3,]
```

>>>>

This script adds plant functional type labels to NEON spectra based on known scientific name. I explored R packages to automate this (BIEN, Taxise, USDA plants) without success, so the code below manually adds labels of interest for *Genus species* names.

- **commonName**
- **pft_MRA** (Plant Functional Type based on Macrosystems proposed labels)
    - coniferous_forest
    - deciduous_forest
    - woody_shrub
    - grass_herb
    bare_ground
- **pft_MODIS** (Plant Functional type based on MODIS land cover classes)
    - evergreen-needleleaf_forest
    - evergreen-broadleaf_forest
    - deciduous-needleleaf_forest
    - deciduous-broadleaf_forest
    - shrub
- **phenology** (evergreen or deciduous)
- **leafType** (broadleaf or needleleaf)
- **growthForm** (shrub or tree, based on NEON field based observations)

I searched for each taxon on USDA plants and determined the phenology and leaf type. Some tips from botany friend Adam Mahood:

- shrubs that are angiosperms are broadleaf
- group = dicot or monocot, you can say broadleaf
- group = gymnosperm will be needleleaf

## Download NEON data

The NEON Woody Vegetation Structure vst_apparentindividual data table contains growthform observations for each individual plant. Since some species can be designated as trees OR shrubs, lets' look at the NEON observations and use whichever growth form is most often observed for each taxon.

@@

@@

# summarize the merged growthform data with "tree" and "shrub" growth forms
cleaned_spectra_merged %>% 
  distinct(spectraID, .keep_all = TRUE) %>% 
  # keep just the tree and shrub growth forms, remove rows with "sampling" and "liana"
  filter(growthForm_simple %in% c("tree","shrub")) %>% 
  group_by(genusSpecies, growthForm_simple) %>% 
    tally() %>% 
  # display as a nicely formatted table
  kableExtra::kable() %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", 
                            full_width = F, 
                            position = "left")
```

@@

@@

## Add PFT label to each spectrum

```{r}

@@

@@
